---
title: "Annotate cell types"
author: "Saskia Freytag"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(SeuratObject)
library(SingleR)
library(ggplot2)
library(SingleCellExperiment)
library(pals)
library(intrinsicDimension)
library(reticulate)
library(readxl)
library(viridis)
library(dplyr)
library(magrittr)
library(BiocParallel)
library(BiocNeighbors)
library(BiocSingular)
library(bluster)
library(scater)
library(scuttle)
library(scran)
library(org.Hs.eg.db)
library(ggrepel)

use_virtualenv("r-reticulate")
scrublet <- import("scrublet")

source("../scripts/plotting.R")

```

```{r}
colours_clusters <- polychrome(36)
names(colours_clusters) <-1:36

colours_labels <- c("MES-like"="#E2726E", "Oligodendrocyte"="#3A4D45", "AC-like"="#8FBAEE", "Mural cell"="#929292", "Mono"="#AB81CB", "TAM-BDM"="#5C9EB7", "Astrocyte"="#5E7BA0", "OPC"="#8FBFAA", "CD4/CD8"="#B13C27", "Neutrophil"="#52A652", "Endothelial"="black", "RG"="#FEFEB0", "Mast"="#A6A2A7", "DC"="#3D8762","TAM-MG"="#3F76B2", "NK"="#E0A23D", "OPC-like"="#B4F0D5", "B cell"="#EABF43", "NPC-like"="#C6A8F6", "Plasma B"="#EBEBEB", "Neuron"="#8470A6")
```

# Set-up data

## Load Seurat 

```{r, eval=FALSE}
spe <- readRDS("../data/raw/SMI-0151_SarahBest_WEHI/6 Analysis/Data objects/seurat_object.Rds")
index_jj98 <- grep("JJ98", colnames(spe), value=TRUE)
```

Isolate JJ98. 

```{r, eval=FALSE}
slide_seq = CreateSeuratObject(counts = spe@assays$Nanostring@counts[, index_jj98], assay="Spatial")
coord_df <- as.data.frame(spe@images$JJ98$centroids@coords)
rownames(coord_df) = colnames(slide_seq)

slide_seq@images$image =  new(
    Class = 'SlideSeq',
    assay = "Spatial",
    key = "image_",
    coordinates = coord_df
  )
```


## Make SingleCellExperiment

We also remove MALAT1 as it drives results. Also remove negative control probes to ensure that these are not taken into account.

```{r, eval=FALSE}
sce_slide <- SingleCellExperiment(list(counts=slide_seq@assays$Spatial@counts),
          colData=as(slide_seq@meta.data, "DataFrame"))

sce_slide <- sce_slide[!rownames(sce_slide) %in% "MALAT1",]
sce_slide <- sce_slide[!grepl("NegPrb", rownames(sce_slide)),]
```

## Do a log normalization

```{r, eval=FALSE}
sce_slide <- logNormCounts(sce_slide)
```

# Analysis for the entire dataset

### Do dimension reduction

We pick 19 dimensions for the PCA. 

```{r, eval=FALSE}
set.seed(0010101010)
sce_slide <- runPCA(sce_slide, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_slide, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_slide, "PCA") <- reducedDim(sce_slide, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_slide <- runUMAP(sce_slide, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 3, min_dist = 0.1)
```

### Clustering

Clustering is performed on the PCA with dimensions set as previously determined and then cluster using 20 nearest neighbors using a Leiden approach.  

```{r, eval=FALSE}
set.seed(0010101010)
clust.leiden <- clusterCells(sce_slide, use.dimred="PCA", 
    BLUSPARAM=SNNGraphParam(k=20, type="rank", cluster.fun="leiden"))
table(clust.leiden)
colData(sce_slide)$Cluster <- clust.leiden

saveRDS(sce_slide, file="../data/sce_jj98.rds")
```

```{r, include=FALSE}
sce_slide <- readRDS("../data/sce_jj98.rds")
```

### Use SingleR to annotate cells

The annotation with the Ruiz-Moreno dataset will be used to annotate cells throughout, but will not be repeated for subsets.

```{r, eval=FALSE}
sc_gbm_big <- readRDS("/stornext/Bioinf/data/lab_brain_cancer/public_datasets/sc_gbm_big/sce_downsampled.rds")
counts(sc_gbm_big) <- logcounts(sc_gbm_big)
rownames(sc_gbm_big) <- rowData(sc_gbm_big)$feature_name

available <- intersect(rownames(sc_gbm_big), rownames(sce_slide))

pred_labels <- SingleR(test=sce_slide[available,], ref=sc_gbm_big[available,], 
    labels=sc_gbm_big$annotation_level_3, de.method="wilcox")
table(pred_labels$labels)

saveRDS(pred_labels, file="../data/prediction.rds")
```

```{r}
pred_labels <- readRDS("../data/prediction.rds")
sce_slide$pred_labels <- pred_labels$labels

tab <- apply(table(sce_slide$Cluster, sce_slide$pred_labels), 1, function(x) x/sum(x))
write.csv(tab, file="../data/prop_per_cluster.csv", quote=FALSE)
```

### Plot UMAPs

```{r}
df <- data.frame(UMAP1=reducedDim(sce_slide, "UMAP")[, 1], UMAP2=reducedDim(sce_slide, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_slide)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Cluster), size=0.455, alpha=0.75) +
  theme_void() + 
  scale_color_manual(values=colours_clusters[1:length(unique(df$Cluster))]) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_clusters.jpg",  bg="white")
```

```{r}
gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=pred_labels), size=0.455, alpha=0.75) +
  theme_void() + scale_color_manual(values=colours_labels) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1
  
ggsave(gg1, file="../figures/UMAP_pred_labels.jpg",  bg="white")  
```

```{r}
make_log_plot(sce_slide, "PTPRC", df)
gg1 <- make_ind_plot(sce_slide, "PTPRC", df)
gg1

ggsave(gg1, file="../figures/PTPRC_ind_all.jpg",  bg="white")
```

```{r}
df$nCount_Spatial <- as.vector(sce_slide$nCount_Spatial)
df$nCount_Feature <- as.vector(sce_slide$nFeature_Spatial)

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=log10(nCount_Spatial))) + geom_point(alpha=0.5, size=0.2) + theme_classic() + scale_color_viridis()

gg1

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=log10(nCount_Feature))) + geom_point(alpha=0.5, size=0.2) + theme_classic() + scale_color_viridis()

gg2
```

## Estimate doublets

We also estimated the number of doublets to ensure that our observations were artefacts of overlapping cells.

```{r}
scrub = scrublet$Scrublet(as.matrix(t(counts(sce_slide))))
pred_doublet = scrub$scrub_doublets()

df$doublet_score <- pred_doublet[[1]]

ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=doublet_score)) + geom_point(alpha=0.5, size=0.2) + theme_classic() + scale_color_viridis()
```

## Annotate cells

```{r}
anno <- read_xlsx("../data/annotation.xlsx")

sce_slide$Anno <- anno$Annotation[match(sce_slide$Cluster, anno$Cluster)]
df$Anno <- sce_slide$Anno

cols_min <- c("Tumour"="indianred", "Immune"="dodgerblue3", "Vasculature"="black")


gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) +
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Anno), size=0.455, alpha=0.75) +
  theme_void() + scale_colour_manual(values=cols_min) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_anno_broad.jpg",  bg="white")
```

```{r}
markers <- findMarkers(sce_slide, groups=sce_slide$Anno, test.type="wilcox", direction="up")

gg1 <- make_log_plot(sce_slide, "CLU", df)
ggsave(gg1, file="../figures/CLU_log_all.jpg",  bg="white")


gg1 <- make_log_plot(sce_slide, "COL4A2", df)
ggsave(gg1, file="../figures/COL4A2_log_all.jpg",  bg="white")


gg1 <- make_log_plot(sce_slide, "CD163", df, order=TRUE)
gg1
ggsave(gg1, file="../figures/CD163_log_all.jpg",  bg="white")
```

# Analyse immune cell only

## Split immune cells

```{r, eval=FALSE}
sce_immune <- sce_slide[, sce_slide$Anno=="Immune"]
```

## First analysis of immune cells

### Dimension reduction

We use 19 dimensions for the PCA in the reclustering of immune cells.

```{r, eval=FALSE}
set.seed(0010101010)
sce_immune <- runPCA(sce_immune, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_immune, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_immune, "PCA") <- reducedDim(sce_immune, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_immune <- runUMAP(sce_immune, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 1, min_dist = 0.01)
```

### Clustering

Clustering is performed on the PCA with dimensions set as previously determined and then cluster using 20 nearest neighbors using a Leiden approach.  

```{r, eval=FALSE}
set.seed(0010101010)
clust.leiden <- clusterCells(sce_immune, use.dimred="PCA", 
    BLUSPARAM=SNNGraphParam(k=12, type="rank", cluster.fun="leiden"))
table(clust.leiden)
colData(sce_immune)$Cluster <- clust.leiden

saveRDS(sce_immune, file="../data/sce_immune_jj98.rds")
```

```{r, include=FALSE}
sce_immune <- readRDS("../data/sce_immune_jj98.rds")
```

### Plot UMAPs

```{r}
df <- data.frame(UMAP1=reducedDim(sce_immune, "UMAP")[, 1], UMAP2=reducedDim(sce_immune, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_immune)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Cluster), size=0.955, alpha=0.75) + theme_void() + scale_color_manual(values=colours_clusters[1:length(unique(df$Cluster))]) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_immune_clusters.jpg",  bg="white")

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=pred_labels), size=0.955, alpha=0.75) + theme_void() + 
  scale_color_manual(values=colours_labels) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg2

ggsave(gg2, file="../figures/UMAP_immune_pred_labels.jpg",  bg="white")
```

```{r, eval=FALSE}
tab <- apply(table(sce_immune$Cluster, sce_immune$pred_labels), 1, function(x) x/sum(x))
write.csv(tab, file="../data/prop_per_cluster_immune.csv", quote=FALSE)
```

```{r}
gg1 <- make_ind_plot(sce_immune, "AHR", df)

ggsave(gg1, file="../figures/AHR_ind_immune.jpg", bg="white", height=5, width=5)
```

### Annotate cells

```{r}
anno <- read_xlsx("../data/annotation_immune.xlsx")

sce_immune$Anno <- anno$Annotation[match(sce_immune$Cluster, anno$Cluster)]
df$Anno <- sce_immune$Anno

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2))  + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Anno), size=0.955, alpha=0.75) +
  theme_void() + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) + 
  scale_color_manual(values=colours_labels) 

gg1

ggsave(gg1, file="../figures/UMAP_immune.jpg", bg="white")
```


## Recluster with only immune cells

Since we identified a population of mesenchymal-like cells in the immune population, we remove these and perform clustering again. For this we use 18 dimensions.

### Dimension reduction

```{r, eval=FALSE}
sce_immune <- sce_immune[, sce_immune$Anno %in% c("CD4/CD8", "DC", "Mono", "TAM-BDM", "TAM-MG") ]

set.seed(0010101010)
sce_immune <- runPCA(sce_immune, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_immune, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_immune, "PCA") <- reducedDim(sce_immune, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_immune <- runUMAP(sce_immune, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 1, min_dist = 0.01)

saveRDS(sce_immune, file="../data/sce_immune_jj98_only.rds")
```

```{r, include=FALSE}
sce_immune <- readRDS("../data/sce_immune_jj98_only.rds")
```

### Plot UMAPs

```{r}
df <- data.frame(UMAP1=reducedDim(sce_immune, "UMAP")[, 1], UMAP2=reducedDim(sce_immune, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_immune)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2))  + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Anno), size=0.955, alpha=0.75) +
  theme_void() + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) + 
  scale_color_manual(values=colours_labels) 

gg1

ggsave(gg1, file="../figures/UMAP_immune_only.jpg", bg="white")
```


```{r}
gg3 <- df %>%
  dplyr::count(Anno, orig.ident) %>%   
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(pct= prop.table(n) * 100) %>%
  ggplot(aes(y=orig.ident, x=pct, fill=Anno)) +
  geom_bar(stat="identity") + scale_fill_manual(values= colours_labels) +
  theme_bw()


gg3
ggsave(gg3, file="../figures/Prop_immune.pdf", height=2, width=7)
```

```{r}
gg1 <- make_ind_plot(sce_immune, "AHR", df)
gg1

ggsave(gg1, file="../figures/AHR_ind_immune.jpg", bg="white")


gg1 <- make_ind_plot(sce_immune, "TGFB1", df)
gg1

ggsave(gg1, file="../figures/TGFB1_ind_immune.jpg", bg="white")
```

# Analysis of tumour cells 

### Split tumour cells

```{r, eval=FALSE}
sce_tumour <- sce_slide[, sce_slide$Anno %in% c("Tumour")]
```

### Dimension reduction

We use 20 dimesions for the PCA of the reclustering of the tumour cells.

```{r, eval=FALSE}
set.seed(0010101010)
sce_tumour <- runPCA(sce_tumour, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_tumour, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_tumour, "PCA") <- reducedDim(sce_tumour, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_tumour <- runUMAP(sce_tumour, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 1, min_dist = 0.1)
```

```{r, eval=FALSE}
set.seed(0010101010)
clust.leiden <- clusterCells(sce_tumour, use.dimred="PCA", 
    BLUSPARAM=SNNGraphParam(k=20, type="rank", cluster.fun="leiden"))
table(clust.leiden)
colData(sce_tumour)$Cluster <- clust.leiden

saveRDS(sce_tumour, file="../data/sce_tumour_jj98.rds")
```

```{r, include=FALSE}
sce_tumour <- readRDS("../data/sce_tumour_jj98.rds")
```

```{r}
df <- data.frame(UMAP1=reducedDim(sce_tumour, "UMAP")[, 1], UMAP2=reducedDim(sce_tumour, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_tumour)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Cluster), size=0.955, alpha=0.75) +
  theme_void() + scale_color_manual(values=colours_clusters[1:length(unique(df$Cluster))]) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_tumour_clusters.jpg",  bg="white")

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=pred_labels), size=0.955, alpha=0.75) +
  theme_void() + scale_color_manual(values=colours_labels) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg2

ggsave(gg2, file="../figures/UMAP_tumour_pred_labels.jpg", bg="white")
```

```{r, eval=FALSE}
tab <- apply(table(sce_tumour$Cluster, sce_tumour$pred_labels), 1, function(x) x/sum(x))
write.csv(tab, file="../data/prop_per_cluster_tumour.csv", quote=FALSE)
```

### Predict cell lables

We will aslo perform an annotation with SingelR with the Couturier dataset set as a reference.

```{r, eval=FALSE}
sc_couturier <- readRDS("/stornext/Bioinf/data/lab_brain_cancer/public_datasets/Deconvolution References/Couturier_dataset_tumor.RDS")
sc_couturier <- logNormCounts(sc_couturier)
rownames(sc_couturier) <- rowData(sc_couturier)$Symbol

available <- intersect(rownames(sc_couturier), rownames(sce_tumour))

pred_labels <- SingleR(test=sce_tumour[available,], ref=sc_couturier[available,], 
    labels=sc_couturier$cluster, de.method="wilcox")
table(pred_labels$labels)

saveRDS(pred_labels, file="../data/prediction_couturier.rds")
```


```{r}
pred_labels <- readRDS("../data/prediction_couturier.rds")

colours_couturier <- c("Mesenchymal"="#E2726E", "Astro"="#8FBAEE", "Progenitor"="#FEFEB0",  "Oligo"="#B4F0D5","Neuronal"="#C6A8F6")

sce_tumour$pred_labels1 <- pred_labels$labels
df$pred_labels1 <- sce_tumour$pred_labels1

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=pred_labels1), size=0.955, alpha=0.75) +
  theme_void() + scale_colour_manual(values=colours_couturier) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1 

ggsave(gg1, file="../figures/UMAP_tumour_couturier.jpg",  bg="white")
  
tab <- apply(table(sce_tumour$Cluster, sce_tumour$pred_labels1), 1, function(x) x/sum(x))
tab
```

### Annotate cells 

```{r}
colours_labels_small <- c("MES-like"="#E2726E", "Oligo"="#336600", "AC-like"="#8FBAEE",
                          "NPC-like/OPC-like/RG"="#E5F9BC", "Mixed"="#FF9933")

anno <- read_xlsx("../data/annotation_tumour.xlsx")

sce_tumour$Anno <- anno$Annotation[match(sce_tumour$Cluster, anno$Cluster)]
df$Anno <- sce_tumour$Anno

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Anno), size=0.955, alpha=0.75) +
  theme_void() + guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) +
  scale_color_manual(values=colours_labels_small) 

gg1

ggsave(gg1, file="../figures/UMAP_tumour.jpg",  bg="white")
```

```{r}
gg1 <- df %>%
  dplyr::count(Anno, orig.ident) %>%       
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(pct= prop.table(n) * 100) %>%
  ggplot(aes(y=orig.ident, x=pct, fill=Anno)) +
  geom_bar(stat="identity") + scale_fill_manual(values= colours_labels_small) +
  theme_bw()

gg1

ggsave(gg1, file="../figures/Prop_tumour.pdf", height=2, width=7)
```

### Differential expression analysis mixed versus everything else

```{r}
sce_tumour$Mixed <- sce_tumour$Anno
sce_tumour$Mixed <- sce_tumour$Mixed=="Mixed"
```

First we run a DEG analysis on mixed versus everything else.

```{r}
markers <- findMarkers(sce_tumour, groups=sce_tumour$Mixed, 
  test.type = c("t"))
  
markers_mixed <- rownames(markers$`TRUE`)[markers$`TRUE`$FDR <= 0.05]

df_volcano <- as.data.frame(markers$`TRUE`)
df_volcano$gene <- rownames(df_volcano)
df_volcano$FDR <- -log10(df_volcano$FDR)

df_volcano_pos <- df_volcano[df_volcano$FDR>(-log10(0.05)) & df_volcano$summary.logFC>0, ]
df_volcano_neg <- df_volcano[df_volcano$FDR>(-log10(0.05)) & df_volcano$summary.logFC<0, ]

gg1 <- ggplot(df_volcano, aes(x=summary.logFC, y=FDR)) + 
  geom_point(colour="black", size=0.5) +
  theme_minimal() + xlab("log fold change") + geom_point(data=df_volcano_pos,
      colour="red") +  geom_point(data=df_volcano_neg, colour="blue") +
  geom_text_repel(data=df_volcano_pos[1:3,], aes(label=gene)) +
   geom_text_repel(data=df_volcano_neg[1:10,], aes(label=gene))

ggsave(gg1, file="../figures/volcano_mixed.jpg")
```

Next, we perform a pathway analysis for these markers and visualize the rank of the genes in the DEG analysis for selected pathways.

```{r}
go_terms <- limma::goana(unique(entrez.ids[markers_mixed]), species="Hs", 
    universe=unique(entrez.ids), trend=TRUE)
go_terms$FDR <- p.adjust(go_terms$P.DE, "fdr")
go_terms <- go_terms[order(go_terms$P.DE), ]
go_terms <- go_terms[go_terms$FDR<0.05, ]

all_gos <- AnnotationDbi::select(org.Hs.eg.db, keys=rownames(go_terms), 
  columns = c('SYMBOL'), keytype = "GOALL")
all_gos <- lapply(split(1:nrow(all_gos), all_gos[,1]), function(x) all_gos[x,4])
all_gos <- lapply(all_gos, function(x) unique(x))
all_gos <- all_gos[rownames(go_terms)]

count_ups_downs <- function(markers_tmp, gene_set){
  
  gene_set <- intersect(rownames(markers_tmp), gene_set)
  tmp <- markers_tmp[gene_set,]
  c(n_down=sum(tmp$summary.logFC<0 & tmp$FDR<0.05) , 
    n_up=sum(tmp$summary.logFC>0 & tmp$FDR<0.05))

}

ups_downs <- lapply(all_gos, function(x) count_ups_downs(markers$`TRUE`, x))
ups_downs <- do.call(rbind,ups_downs)
go_terms <- cbind(go_terms, ups_downs)
go_terms <- go_terms[go_terms$N>10,]

go_terms <- go_terms[order(go_terms$n_up/go_terms$DE, decreasing=TRUE),]
write.table(go_terms, file="../data/go_terms_mixed.tsv", sep="\t", quote=FALSE, col.names=TRUE, row.names = TRUE)
```

```{r}
selected_terms <- c("GO:0007399", "GO:0022008", "GO:0009888", "GO:0030182",
        "GO:0071363", "GO:0048699", "GO:0007049", "GO:0005929", "GO:0048666")
signs <- sign(markers$`TRUE`$summary.logFC) 
markers$`TRUE`$rank <- log10(markers$`TRUE`$FDR)*-1
markers$`TRUE`$rank <- markers$`TRUE`$rank*signs
ranks <- rank(markers$`TRUE`$rank)
max_rank <- max(ranks[markers$`TRUE`$FDR < 0.05 & 
  markers$`TRUE`$summary.logFC < 0])
min_rank <- min(ranks[markers$`TRUE`$FDR < 0.05 & 
  markers$`TRUE`$summary.logFC > 0])

plot_df <- lapply(selected_terms, function(x) find_ranks(markers$`TRUE`, all_gos, x))
plot_df <- do.call(rbind, plot_df)
plot_df$Term <- go_terms$Term[match(plot_df$GO, rownames(go_terms))]
plot_df$Term <- paste0(plot_df$Term, " (" , round(go_terms$FDR[match(plot_df$GO, rownames(go_terms))], 4), ")")

gg1 <- ggplot(plot_df, aes(y=Term, x=Rank)) + 
  geom_jitter(alpha=0.5, size=0.2, height=0.2) +       
  theme_minimal() + annotate("rect", xmin = 0, xmax = max_rank, 
      ymin = 0, ymax = length(selected_terms)+0.5, alpha = .1, fill = "blue") +
  annotate("rect", 
      xmin = min_rank, xmax = 1000, ymin = 0, ymax = length(selected_terms)+0.5, 
      alpha = .1, fill = "red") + theme(axis.text.x= element_blank(),
      axis.title.x = element_blank(), axis.title.y=element_blank())

ggsave(gg1, file="../figures/GO_Terms_mixed.pdf")
```

### Find markers for the mixed population

```{r}
markers <- scoreMarkers(sce_tumour, sce_tumour$Anno)

markers_mixed <- markers$`Mixed`
markers_mixed <- rownames(markers_mixed)[markers_mixed$mean.AUC>0.5]

go_terms <- limma::goana(unique(entrez.ids[markers_mixed]), species="Hs", 
    universe=unique(entrez.ids), trend=TRUE)
go_terms$FDR <- p.adjust(go_terms$P.DE, "fdr")
go_terms <- go_terms[order(go_terms$P.DE), ]
go_terms <- go_terms[go_terms$FDR<0.05, ]

write.table(go_terms, file="../data/go_terms_up_mixed.tsv", sep="\t", 
quote=FALSE, col.names=TRUE, row.names = TRUE)

kegg_terms <- limma::kegga(unique(entrez.ids[markers_mixed]), species="Hs", 
    universe=unique(entrez.ids), trend=TRUE)
kegg_terms$FDR <- p.adjust(kegg_terms$P.DE, "fdr")
kegg_terms <- kegg_terms[order(kegg_terms$P.DE), ]
kegg_terms <- go_terms[go_terms$FDR<0.05, ]

write.table(kegg_terms, file="../data/kegg_terms_up_mixed.tsv", sep="\t", 
quote=FALSE, col.names=TRUE, row.names = TRUE)

gg1 <- plotGroupedHeatmap(sce_tumour, features=c("MZT2A", "NPPC", 
"WIF1", "KRT16", "IGF2", "MEG3"), group="Anno", 
    center=TRUE)
    
ggsave(gg1, file="../figures/mixed_heatmap_markers.pdf", height=5, width=7)    

```

### Estimate the cell phase of the tumour cells

```{r, eval=FALSE}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

s.genes <- intersect(rownames(sce_tumour), s.genes)
s.genes 
g2m.genes <- intersect(rownames(sce_tumour), g2m.genes)
g2m.genes 

sce_tumour_seurat <- as.Seurat(sce_tumour)

cell_cycle <- CellCycleScoring(sce_tumour_seurat, 
    s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

df$Sphase <- cell_cycle$S.Score
df$G2Mphase <- cell_cycle$G2M.Score

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=Sphase)) + geom_point(alpha=1, size=0.5) + theme_classic() + scale_color_viridis()
gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=G2Mphase)) + geom_point(alpha=1, size=0.5) + theme_classic() + scale_color_viridis()

ggsave(gg1, file="../figures/Sphase_tumour.jpg",  bg="white")
ggsave(gg1, file="../figures/G2Mphase_tumour.jpg",  bg="white")
```

# Save all annotations

```{r, eval=FALSE}
anno_df <- data.frame(Barcode=colnames(sce_slide), Annotation=sce_slide$Anno)
index_immune <- match(colnames(sce_immune), anno_df$Barcode)
anno_df$Annotation[index_immune] <- sce_immune$Anno
index_tumour <- match(colnames(sce_tumour), anno_df$Barcode)
anno_df$Annotation[index_tumour] <- sce_tumour$Anno
anno_df$Annotation[anno_df$Annotation=="Immune"] <- "MES-like"

colours <- c("MES-like"="#DD716E", "TAM-MG"="#3C77B2", "AC-like"="#91B8E1", 
             "TAM-BDM"="#5A9EB5",  "NPC-like/OPC-like/RG"="#E5F9BC", "Mixed"="#FF9933", "Oligo"="#336600", "Mono"="#A280B4",
             "Vasculature"="black", "CD4/CD8"="#BB5C3C")

sce_slide$Anno_final <- anno_df$Annotation

df <- data.frame(UMAP1=reducedDim(sce_slide, "UMAP")[, 1], UMAP2=reducedDim(sce_slide, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_slide)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Anno_final), size=0.455, alpha=0.75) +
  theme_void() + 
  scale_color_manual(values=colours) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

ggsave(gg1, file="../figures/UMAP_full_anno.jpg", bg="white")


saveRDS(anno_df, file="../data/annotation_all.rds")
```