---
title: "Annotate cell types"
author: "Saskia Freytag"
date: "2023-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(SeuratObject)
library(SingleR)
library(ggplot2)
library(SingleCellExperiment)
library(pals)
library(intrinsicDimension)
library(reticulate)
library(readxl)
library(viridis)
library(dplyr)
library(magrittr)
library(BiocParallel)
library(BiocNeighbors)
library(BiocSingular)
library(bluster)
library(scater)
library(scuttle)
library(scran)
library(org.Hs.eg.db)
library(ggrepel)
library(readxl)

use_virtualenv("r-reticulate")
scrublet <- import("scrublet")

source("../scripts/plotting.R")

```

```{r}
colours_clusters <- polychrome(36)
names(colours_clusters) <-1:36

colours_labels <- c("MES-like"="#E2726E", "Oligodendrocyte"="#3A4D45", "AC-like"="#8FBAEE", "Mural cell"="#929292", "Mono"="#AB81CB", "TAM-BDM"="#5C9EB7", "Astrocyte"="#5E7BA0", "OPC"="#8FBFAA", "CD4/CD8"="#B13C27", "Neutrophil"="#52A652", "Endothelial"="black", "RG"="#FEFEB0", "Mast"="#A6A2A7", "DC"="#3D8762","TAM-MG"="#3F76B2", "NK"="#E0A23D", "OPC-like"="#B4F0D5", "B cell"="#EABF43", "NPC-like"="#C6A8F6", "Plasma B"="#EBEBEB", "Neuron"="#8470A6")

deconvolution_markers <- read_xlsx("../data/processed/deconvolution_signatures.xlsx")
```

# Set-up data

## Load Seurat 

```{r, eval=FALSE}
spe <- readRDS("../data/raw/SMI-0151_SarahBest_WEHI/6 Analysis/Data objects/seurat_object.Rds")
index_jj98 <- grep("JJ98", colnames(spe), value=TRUE)
```


## Make SingleCellExperiment

We also remove MALAT1 as it drives results. Also remove negative control probes to ensure that these are not taken into account.

We will also filter any cells that have less than 20 different detected genes. 

```{r, eval=FALSE}
sce_slide <- SingleCellExperiment(list(counts=spe@assays$Nanostring@counts[, index_jj98]),
          colData=as(spe@meta.data[index_jj98,], "DataFrame"))

sce_slide <- sce_slide[!rownames(sce_slide) %in% "MALAT1",]
sce_slide <- sce_slide[!grepl("NegPrb", rownames(sce_slide)),]

sce_slide <- sce_slide[,!sce_slide$nFeature_Nanostring<20]
```

## Do a log normalization

```{r, eval=FALSE}
sce_slide <- logNormCounts(sce_slide)
```

# Analysis for the entire dataset

## Do dimension reduction

We pick 19 dimensions for the PCA. 

```{r, eval=FALSE}
set.seed(0010101010)
sce_slide <- runPCA(sce_slide, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_slide, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_slide, "PCA") <- reducedDim(sce_slide, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_slide <- runUMAP(sce_slide, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 3, min_dist = 0.1)
```

## Clustering

Clustering is performed on the PCA with dimensions set as previously determined and then cluster using 20 nearest neighbors using a Leiden approach.  

```{r, eval=FALSE}
set.seed(0010101010)
clust.leiden <- clusterCells(sce_slide, use.dimred="PCA", 
    BLUSPARAM=SNNGraphParam(k=20, type="rank", cluster.fun="leiden"))
table(clust.leiden)
colData(sce_slide)$Cluster <- clust.leiden

saveRDS(sce_slide, file="../data/processed/sce_jj98.rds")
```

```{r, include=FALSE}
sce_slide <- readRDS("../data/processed/sce_jj98.rds")
```

## Investigate diversity of genes

```{r}
df <- data.frame(UMAP1=reducedDim(sce_slide, "UMAP")[, 1], UMAP2=reducedDim(sce_slide, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_slide)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=log10(nCount_Nanostring)), size=0.455, alpha=0.75) +
  theme_void() + 
  scale_color_viridis_c()

gg1

ggsave(gg1, file="../figures/UMAP_counts.jpg",  bg="white")
 
gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=log10(nFeature_Nanostring)), size=0.455, alpha=0.75) +
  theme_void() + 
  scale_color_viridis_c()

gg2

ggsave(gg2, file="../figures/UMAP_features.jpg",  bg="white")

```

```{r}
deconvolution_markers$In <- deconvolution_markers$Gene %in% rownames(sce_slide)

deconvolution_markers %>% group_by(Population) %>% summarise(sum_markers=sum(In))
```


## Use SingleR to annotate cells

The annotation with the Ruiz-Moreno dataset will be used to annotate cells throughout, but will not be repeated for subsets.

```{r, eval=FALSE}
sc_gbm_big <- readRDS("/stornext/Bioinf/data/lab_brain_cancer/public_datasets/sc_gbm_big/sce_downsampled.rds")
counts(sc_gbm_big) <- logcounts(sc_gbm_big)
rownames(sc_gbm_big) <- rowData(sc_gbm_big)$feature_name

available <- intersect(rownames(sc_gbm_big), rownames(sce_slide))

pred_labels <- SingleR(test=sce_slide[available,], ref=sc_gbm_big[available,], 
    labels=sc_gbm_big$annotation_level_3, de.method="wilcox", num.threads = 6)
table(pred_labels$labels)

saveRDS(pred_labels, file="../data/processed/prediction.rds")
```

```{r}
pred_labels <- readRDS("../data/processed/prediction.rds")
sce_slide$pred_labels <- pred_labels$labels

tab <- apply(table(sce_slide$Cluster, sce_slide$pred_labels), 1, function(x) x/sum(x))
write.csv(tab, file="../data/processed/prop_per_cluster.csv", quote=FALSE)
```

### Plot UMAPs

```{r}
df <- data.frame(UMAP1=reducedDim(sce_slide, "UMAP")[, 1], UMAP2=reducedDim(sce_slide, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_slide)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Cluster), size=0.455, alpha=0.75) +
  theme_void() + 
  scale_color_manual(values=colours_clusters[1:length(unique(df$Cluster))]) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_clusters.jpg",  bg="white")
```

```{r}
gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=pred_labels), size=0.455, alpha=0.75) +
  theme_void() + scale_color_manual(values=colours_labels) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1
  
ggsave(gg1, file="../figures/UMAP_pred_labels.jpg",  bg="white")  
```

```{r}
make_log_plot(sce_slide, "PTPRC", df)
gg1 <- make_ind_plot(sce_slide, "PTPRC", df)
gg1

ggsave(gg1, file="../figures/PTPRC_ind_all.jpg",  bg="white")

gg1 <- make_log_plot(sce_slide, "CD163", df)
gg1

ggsave(gg1, file="../figures/CD163_all.jpg",  bg="white")

gg1 <- make_log_plot(sce_slide, "CLU", df)
gg1

ggsave(gg1, file="../figures/CLU_all.jpg",  bg="white")

gg1 <- make_log_plot(sce_slide, "COL4A2", df)
gg1

ggsave(gg1, file="../figures/COL4A2_all.jpg",  bg="white")

```

## Final annotation

```{r}
anno <- read_xlsx("../data/processed/annotation_broad.xlsx")

sce_slide$Anno <- anno$Annotation[match(sce_slide$Cluster, anno$Cluster)]
df$Anno <- sce_slide$Anno

cols_min <- c("Tumour"="indianred", "Immune"="dodgerblue3", "Vasculature"="black", "Oligodendrocyte"="#3A4D45")


gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) +
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Anno), size=0.455, alpha=0.75) +
  theme_void() + scale_colour_manual(values=cols_min) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_anno_broad.jpg",  bg="white")

```

# Tumour only annotation

## Split tumour cells

```{r, eval=FALSE}
sce_tumour <- sce_slide[, sce_slide$Anno %in% c("Tumour")]
```

```{r, include=FALSE}
sce_tumour <- readRDS("../data/processed/sce_tumour_jj98.rds")
```

## Check gene diversity

```{r}
table(colSums(as.matrix(counts(sce_tumour[rownames(sce_tumour) %in% deconvolution_markers$Gene,])))>0)/ncol(sce_tumour)*100
```

## Dimension reduction

We use 20 dimesions for the PCA of the reclustering of the tumour cells.

```{r, eval=FALSE}
set.seed(0010101010)
sce_tumour <- runPCA(sce_tumour, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_tumour, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_tumour, "PCA") <- reducedDim(sce_tumour, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_tumour <- runUMAP(sce_tumour, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 1, min_dist = 0.1)
```

```{r, eval=FALSE}
set.seed(0010101010)
clust.leiden <- clusterCells(sce_tumour, use.dimred="PCA", 
    BLUSPARAM=SNNGraphParam(k=20, type="rank", cluster.fun="leiden"))
table(clust.leiden)
colData(sce_tumour)$Cluster <- clust.leiden

saveRDS(sce_tumour, file="../data/processed/sce_tumour_jj98.rds")
```


```{r}
df <- data.frame(UMAP1=reducedDim(sce_tumour, "UMAP")[, 1], UMAP2=reducedDim(sce_tumour, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_tumour)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Cluster), size=0.955, alpha=0.75) +
  theme_void() + scale_color_manual(values=colours_clusters[1:length(unique(df$Cluster))]) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_tumour_clusters.jpg",  bg="white")

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=pred_labels), size=0.955, alpha=0.75) +
  theme_void() + scale_color_manual(values=colours_labels) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg2

ggsave(gg2, file="../figures/UMAP_tumour_pred_labels.jpg", bg="white")
```

```{r, eval=FALSE}
tab <- apply(table(sce_tumour$Cluster, sce_tumour$pred_labels), 1, function(x) x/sum(x))
write.csv(tab, file="../data/processed/prop_per_cluster_tumour.csv", quote=FALSE)
```

### Predict cell lables

We will also perform an annotation with SingelR with the Couturier dataset set as a reference.

```{r, eval=FALSE}
sc_couturier <- readRDS("/stornext/Bioinf/data/lab_brain_cancer/public_datasets/Deconvolution References/Couturier_dataset_tumor.RDS")
sc_couturier <- logNormCounts(sc_couturier)
rownames(sc_couturier) <- rowData(sc_couturier)$Symbol

available <- intersect(rownames(sc_couturier), rownames(sce_tumour))

pred_labels <- SingleR(test=sce_tumour[available,], ref=sc_couturier[available,], 
    labels=sc_couturier$cluster, de.method="wilcox", num.threads=6)
table(pred_labels$labels)

saveRDS(pred_labels, file="../data/processed/prediction_couturier.rds")
```


```{r}
pred_labels <- readRDS("../data/processed/prediction_couturier.rds")

colours_couturier <- c("Mesenchymal"="#E2726E", "Astro"="#8FBAEE", "Progenitor"="#FEFEB0",  "Oligo"="#B4F0D5","Neuronal"="#C6A8F6")

sce_tumour$pred_labels1 <- pred_labels$labels
df$pred_labels1 <- sce_tumour$pred_labels1

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=pred_labels1), size=0.955, alpha=0.75) +
  theme_void() + scale_colour_manual(values=colours_couturier) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1 

ggsave(gg1, file="../figures/UMAP_tumour_couturier.jpg",  bg="white")
  
tab <- apply(table(sce_tumour$Cluster, sce_tumour$pred_labels1), 1, function(x) x/sum(x))
write.csv(tab, file="../data/processed/prop_per_cluster_tumour_couturier.csv", quote=FALSE)
```

## TOP2A

```{r}
make_log_plot(sce_tumour, "TOP2A", df)
gg1 <- make_log_plot(sce_tumour, "TOP2A", df)
gg1

ggsave(gg1, file="../figures/TOP2A_log_tumour.jpg",  bg="white")

gg1 <- make_log_plot(sce_tumour, "APOD", df)
gg1

ggsave(gg1, file="../figures/APOD_log_tumour.jpg",  bg="white")


gg1 <- make_log_plot(sce_tumour, "NTRK2", df)
gg1

ggsave(gg1, file="../figures/NTRK2_log_tumour.jpg",  bg="white")

gg1 <- make_log_plot(sce_tumour, "CHI3L1", df)
gg1

ggsave(gg1, file="../figures/CHI3L1_log_tumour.jpg",  bg="white")


gg1 <- make_log_plot(sce_tumour, "STMN1", df)
gg1

ggsave(gg1, file="../figures/STMN1_log_tumour.jpg",  bg="white")

```

## Final annotation

```{r}
anno <- read_xlsx("../data/processed/annotation_tumour.xlsx")

sce_tumour$Anno <- anno$Annotation[match(sce_tumour$Cluster, anno$Cluster)]
df$Anno <- sce_tumour$Anno

colours_new <- c("Mesenchymal"="#E2726E", "Astro"="#8FBAEE", "Progenitor/Neuronal/Oligo"="#E5F9BC")

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) +
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Anno), size=0.455, alpha=0.75) +
  theme_void() + scale_colour_manual(values=colours_new) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_anno_tumour.jpg",  bg="white")

```

```{r}
gg1 <- df %>%
  dplyr::count(Anno, orig.ident) %>%       
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(pct= prop.table(n) * 100) %>%
  ggplot(aes(y=orig.ident, x=pct, fill=Anno)) +
  geom_bar(stat="identity") + scale_fill_manual(values= colours_new) +
  theme_bw()

gg1

ggsave(gg1, file="../figures/Prop_tumour.pdf", height=2, width=7)
```

```{r}
df$TOP2A <- as.vector(t(counts(sce_tumour["TOP2A",])>0))
  
gg1 <- df %>%
  dplyr::count(Anno, TOP2A) %>%       
  dplyr::group_by(Anno) %>%
  dplyr::mutate(pct= prop.table(n) * 100) %>%
  dplyr::filter(TOP2A) %>% 
  ggplot(aes(y=pct, x=Anno, fill=Anno)) +
  geom_bar(stat="identity", colour="black") + scale_fill_manual(values= colours_new) +
  theme_bw()

gg1

ggsave(gg1, file="../figures/Prop_TOP2A_tumour.pdf", height=2, width=7)
```

# Immune only annotation

```{r, eval=FALSE}
sce_immune <- sce_slide[, sce_slide$Anno=="Immune"]
```

## Dimension reduction

We use 19 dimensions for the PCA in the reclustering of immune cells.

```{r, eval=FALSE}
set.seed(0010101010)
sce_immune <- runPCA(sce_immune, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_immune, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_immune, "PCA") <- reducedDim(sce_immune, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_immune <- runUMAP(sce_immune, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 1, min_dist = 0.01)
```

## Clustering

Clustering is performed on the PCA with dimensions set as previously determined and then cluster using 12 nearest neighbors using a Leiden approach.  

```{r, eval=FALSE}
set.seed(0010101010)
clust.leiden <- clusterCells(sce_immune, use.dimred="PCA", 
    BLUSPARAM=SNNGraphParam(k=12, type="rank", cluster.fun="leiden"))
table(clust.leiden)
colData(sce_immune)$Cluster <- clust.leiden

saveRDS(sce_immune, file="../data/processed/sce_immune_jj98.rds")
```

```{r, include=FALSE}
sce_immune <- readRDS("../data/processed/sce_immune_jj98.rds")
```

### Plot UMAPs

```{r}
df <- data.frame(UMAP1=reducedDim(sce_immune, "UMAP")[, 1], UMAP2=reducedDim(sce_immune, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_immune)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Cluster), size=0.955, alpha=0.75) + theme_void() + scale_color_manual(values=colours_clusters[1:length(unique(df$Cluster))]) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_immune_clusters.jpg",  bg="white")

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=pred_labels), size=0.955, alpha=0.75) + theme_void() + 
  scale_color_manual(values=colours_labels) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg2

ggsave(gg2, file="../figures/UMAP_immune_pred_labels.jpg",  bg="white")
```

```{r, eval=FALSE}
tab <- apply(table(sce_immune$Cluster, sce_immune$pred_labels), 1, function(x) x/sum(x))
write.csv(tab, file="../data/processed/prop_per_cluster_immune.csv", quote=FALSE)
```

```{r}
gg1 <- make_ind_plot(sce_immune, "AHR", df)

gg1

ggsave(gg1, file="../figures/AHR_ind_immune.jpg", bg="white", height=5, width=5)
```
## Annotate cells

```{r}
anno <- read_xlsx("../data/processed/annotation_immune.xlsx")

sce_immune$Anno <- anno$Annotation[match(sce_immune$Cluster, anno$Cluster)]
df$Anno <- sce_immune$Anno

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2))  + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Anno), size=0.955, alpha=0.75) +
  theme_void() + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) + 
  scale_color_manual(values=colours_labels) 

gg1

ggsave(gg1, file="../figures/UMAP_immune.jpg", bg="white")
```
## Recluster with only immune cells

Since we identified a population of mesenchymal-like cells in the immune population, we remove these and perform clustering again. For this we use 18 dimensions.

### Dimension reduction

```{r, eval=FALSE}
sce_immune <- sce_immune[, sce_immune$Anno %in% c("CD4/CD8", "DC", "Mono", "TAM-BDM", "TAM-MG") ]

set.seed(0010101010)
sce_immune <- runPCA(sce_immune, ncomponents=50,
    BSPARAM=BiocSingular::IrlbaParam(),  BPPARAM=MulticoreParam(10))

dim_est <- maxLikGlobalDimEst(reducedDim(sce_immune, "PCA"), k=20, unbiased=TRUE)
dim_est <- ceiling(dim_est$dim.est)
print(dim_est)

reducedDim(sce_immune, "PCA") <- reducedDim(sce_immune, "PCA")[,1:dim_est] 

set.seed(0010101010)
sce_immune <- runUMAP(sce_immune, dimred="PCA", 
   BPPARAM=MulticoreParam(10), spread = 1, min_dist = 0.01)

saveRDS(sce_immune, file="../data/processed/sce_immune_jj98_only.rds")
```

```{r, include=FALSE}
sce_immune <- readRDS("../data/processed/sce_immune_jj98_only.rds")
```

### Plot UMAPs

```{r}
df <- data.frame(UMAP1=reducedDim(sce_immune, "UMAP")[, 1], UMAP2=reducedDim(sce_immune, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_immune)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2))  + 
  geom_point(col="black", size=1.5) +
  geom_point(col="white", size=1) +
  geom_point(aes(col=Anno), size=0.955, alpha=0.75) +
  theme_void() + 
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3))) + 
  scale_color_manual(values=colours_labels) 

gg1

ggsave(gg1, file="../figures/UMAP_immune_only.jpg", bg="white")
```


```{r}
gg3 <- df %>%
  dplyr::count(Anno, orig.ident) %>%   
  dplyr::group_by(orig.ident) %>%
  dplyr::mutate(pct= prop.table(n) * 100) %>%
  ggplot(aes(y=orig.ident, x=pct, fill=Anno)) +
  geom_bar(stat="identity", colour="black") + scale_fill_manual(values= colours_labels) +
  theme_bw()


gg3
ggsave(gg3, file="../figures/Prop_immune.pdf", height=2, width=7)
```

```{r}
gg1 <- make_ind_plot(sce_immune, "AHR", df)
gg1

ggsave(gg1, file="../figures/AHR_ind_immune.jpg", bg="white")
```

# Putting annotation together

```{r}
anno_df <- data.frame(Barcode=colnames(sce_slide), Annotation=sce_slide$Anno)
index_immune <- match(colnames(sce_immune), anno_df$Barcode)
anno_df$Annotation[index_immune] <- sce_immune$Anno
index_tumour <- match(colnames(sce_tumour), anno_df$Barcode)
anno_df$Annotation[index_tumour] <- sce_tumour$Anno
anno_df$Annotation[anno_df$Annotation=="Immune"] <- "Mesenchymal"

colours <- c("Mesenchymal"="#E2726E", "Astro"="#8FBAEE", "Progenitor/Neuronal/Oligo"="#E5F9BC",
             "Vasculature"="black",  "Oligodendrocyte"="#3A4D45", "Mono"="#AB81CB", "TAM-BDM"="#5C9EB7",  "CD4/CD8"="#B13C27","TAM-MG"="#3F76B2")

sce_slide$Anno_final <- anno_df$Annotation

df <- data.frame(UMAP1=reducedDim(sce_slide, "UMAP")[, 1], UMAP2=reducedDim(sce_slide, "UMAP")[, 2])
df <- as.data.frame(cbind(df, colData(sce_slide)))

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2)) + 
  geom_point(col="black", size=0.9) +
  geom_point(col="white", size=0.5) +
  geom_point(aes(col=Anno_final), size=0.455, alpha=0.75) +
  theme_void() + 
  scale_color_manual(values=colours) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=3)))

gg1

ggsave(gg1, file="../figures/UMAP_full_anno.jpg", bg="white")

saveRDS(sce_slide, file="../data/processed/sce_jj98_final.rds")

saveRDS(anno_df, file="../data/annotation_all.rds")
```

## Investigate cell cycle

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

s.genes <- intersect(rownames(sce_slide), s.genes)
s.genes 
g2m.genes <- intersect(rownames(sce_slide), g2m.genes)
g2m.genes 

sce_tumour_seurat <- as.Seurat(sce_slide)

cell_cycle <- CellCycleScoring(sce_tumour_seurat, 
    s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

df$Sphase <- cell_cycle$S.Score
df$G2Mphase <- cell_cycle$G2M.Score

gg1 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=Sphase)) + geom_point(alpha=1, size=0.5) + theme_classic() + scale_color_viridis()

gg1

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=G2Mphase)) + geom_point(alpha=1, size=0.5) + theme_classic() + scale_color_viridis()

gg2

ggsave(gg1, file="../figures/Sphase_tumour.jpg",  bg="white")
ggsave(gg1, file="../figures/G2Mphase_tumour.jpg",  bg="white")
```

## Investigate hypoxia

```{r}
tab <- AnnotationDbi::select(org.Hs.eg.db, keytype="SYMBOL", keys=rownames(sce_slide), columns="GOALL")
by.go <- split(tab[,1], tab[,2])

by_go_hypoxia <- by.go["GO:0001666"]

aggregated <- sumCountsAcrossFeatures(sce_slide, by_go_hypoxia,
    exprs_values="logcounts", average=FALSE)

df$hypoxia <- aggregated[1,]

gg2 <- ggplot(df[sample(1:nrow(df)),], aes(x=UMAP1, y=UMAP2, col=hypoxia)) + geom_point(alpha=1, size=0.5) + theme_classic() + scale_color_viridis() 

gg2

ggsave(gg2, file="../figures/hypoxia.jpg",  bg="white")

df %>% group_by(Anno_final, fov) %>% summarize(average_hypoxia=mean(hypoxia))
```




